---
title: "Data Analysis Project"
author: "BIOST 536"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

options(digits = 3) ## Formats output to 3 digits
library(lmtest)
library(tidyverse)
library(kableExtra)
load("tbdata.Rdata")
tbdata = tibble(tbdata)
```

In order to make the household density variable that we are interested in in order to perform a grouped linear adjustment with ordinal variables, we must mutate our data by dividing family size by number of rooms.

```{r}
tbdata = tbdata %>% mutate(famSize = case_when(
  famSize == "<4" ~ 1,
  famSize == "4-6" ~ 2,
  famSize == ">6" ~ 3
))


tbdata = tbdata %>% mutate(roomNo = case_when(
  roomNo == "1" ~ 1,
  roomNo == "2" ~ 2,
  roomNo == "3+" ~ 3
))

tbdata = tbdata %>% mutate(householdDensity = (famSize / roomNo))
```

Now, we can fit a logistic regression model to estimate the risk of TB infection associated with the number of windows in a household, adjusting for sex and our household density variable.

```{r}
tb_mod_glm = glm(case ~ windowNo + male + householdDensity,
                 family=binomial, 
                 data=tbdata)

summary(tb_mod_glm)

tb_mod_glm_null = glm(case ~ male + householdDensity,
                      family=binomial,
                      data=tbdata)
```

In this case, the model estimates that the odds of TB infection for a subpopulation of individuals with 1 window in their house, with the same sex and household density, is `r exp(coef(tb_mod_glm))["windowNo1"]` times the odds of TB infection for a subpopulation of individuals without windows in their house with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo1", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo1", "97.5 %"])`].

Comparing individuals with 2-3 windows vs no windows, the OR is now `r exp(coef(tb_mod_glm))["windowNo2-3"]` with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo2-3", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo2-3", "97.5 %"])`].

Comparing individuals with 4 or more windows vs no windows, the OR is now `r exp(coef(tb_mod_glm))["windowNo4+"]` with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo4+", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo4+", "97.5 %"])`].

Since we have full and null models at this point, we can also run an LRtest in order to assess the significance of the number of windows in a house being a risk factor for TB infection. We'll use the significance level $\alpha = 0.05$

```{r}
lrtest(tb_mod_glm, tb_mod_glm_null)
```

At the aforementioned significance level, we can reject the null hypothesis that the number of windows in a house does not have an effect on the risk of TB infection for this population of individuals.
