---
title: "Data Analysis Project"
author: "Group 1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

options(digits = 3) ## Formats output to 3 digits
library(lmtest)
library(tidyverse)
library(kableExtra)
load("tbdata.Rdata")
tbdata = tibble(tbdata)
```

```{r}
# mutate data to perform grouped linear adjustment with ordinal variables
tbdata = tbdata %>% mutate(famSize = case_when(
  famSize == "<4" ~ 1,
  famSize == "4-6" ~ 2,
  famSize == ">6" ~ 3
))


tbdata = tbdata %>% mutate(roomNo = case_when(
  roomNo == "1" ~ 1,
  roomNo == "2" ~ 2,
  roomNo == "3+" ~ 3
))

# divide ordinal familySize by roomNo to create crude householdDensity covariate
tbdata = tbdata %>% mutate(householdDensity = (famSize / roomNo))

# fit logreg model to estimate risk of TB associated with windows, adjusting for sex, household density, and income
tb_mod_glm = glm(case ~ windowNo + male + householdDensity + income,
                 family=binomial, 
                 data=tbdata)

summary(tb_mod_glm)

tb_mod_glm_null = glm(case ~ male + householdDensity + income,
                      family=binomial,
                      data=tbdata)
```

In this case, the model estimates that the odds of TB infection for a subpopulation of individuals with 1 window in their house, with the same sex and household density, is `r exp(coef(tb_mod_glm))["windowNo1"]` times the odds of TB infection for a subpopulation of individuals without windows in their house with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo1", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo1", "97.5 %"])`].

Comparing individuals with 2-3 windows vs no windows, the OR is now `r exp(coef(tb_mod_glm))["windowNo2-3"]` with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo2-3", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo2-3", "97.5 %"])`].

Comparing individuals with 4 or more windows vs no windows, the OR is now `r exp(coef(tb_mod_glm))["windowNo4+"]` with a 95% CI using model based SEs of [`r exp(confint(tb_mod_glm)["windowNo4+", "2.5 %"])`, `r exp(confint(tb_mod_glm)["windowNo4+", "97.5 %"])`].

Since we have full and null models at this point, we can also run an LRtest in order to assess the significance of the number of windows in a house being a risk factor for TB infection. We'll use the significance level $\alpha = 0.05$

```{r}
# perform LR test on full and null models to assess significance of numWindows
lrtest(tb_mod_glm, tb_mod_glm_null)
```

At the aforementioned significance level, we can reject the null hypothesis that the number of windows in a house does not have an effect on the risk of TB infection for this population of individuals.


```{r Table1, eval=F}
# Set environment and load data
rm(list=ls())
library(tidyverse, warn.conflicts = F)
library(gt)
library(gtable)
library(gtsummary)

setwd('/Users/calenmendall/Desktop/UW_school_projects/Biost536/BIOST536_Extended_abstract/')
load(file = 'tbdata.RData')

## Add in household density measure
tbdata = tbdata %>% mutate(famSize = case_when(
  famSize == "<4" ~ 1,
  famSize == "4-6" ~ 2,
  famSize == ">6" ~ 3
))

tbdata = tbdata %>% mutate(roomNo = case_when(
  roomNo == "1" ~ 1,
  roomNo == "2" ~ 2,
  roomNo == "3+" ~ 3
))

tbdata = tbdata %>% mutate(householdDensity = round((famSize / roomNo),2))

## Select relevant subset of data to table
tbdata_sub = tbdata %>% select(
  'Case'= case,
  'Number of Windows in home' = windowNo,
  "Sex" = male,
  'Age'= age,
  "Income" = income,
  "Family Size" = famSize, 
  "Number of Rooms" = roomNo,
  'Household density' = householdDensity, 
  "BCG Scar" = BCGscar,
  "Household member with TB" = houseMemberTb
)

## Some transformations for the table
tbdata_sub$Sex = factor(tbdata_sub$Sex, 
                        labels = c('Female', 'Male'))
tbdata_sub$Case = factor(tbdata_sub$Case,
                         labels = c('Control', 'Case'))
tbdata_sub$Income = factor(tbdata_sub$Income, 
                           labels= c('Low', 'Mid', 'High'),
                           levels = c('low','mid','high'))
tbdata_sub$`Number of Rooms` = factor(tbdata_sub$`Number of Rooms`,
                                      labels= c('1', '2', '3+'),
                                      levels = c(1, 2, 3))
tbdata_sub$`Family Size` = factor(tbdata_sub$`Family Size`, 
                                  labels= c('<4', '4-6', '>6'),
                                  levels = c(1, 2, 3))
tbdata_sub$`BCG Scar` = factor(tbdata_sub$`BCG Scar`, 
                               labels = c('No', 'Yes'))
tbdata_sub$`Household member with TB`= factor(tbdata_sub$`Household member with TB`, 
    labels = c('No', 'Yes'))

## Generating the table and formatting
tab = tbl_summary(tbdata_sub, by = 'Case') %>% add_overall(last=T)%>% as_gt()

tab = tab %>% tab_spanner('Tuberculosis infection status', columns = c(6,7,8), gather = T, level = 1)

tab2 = tab %>% tab_options(table.font.size = 14, heading.title.font.size = 15,
                    row_group.font.weight = 'bold', 
                    column_labels.font.weight = 'bold', 
                    row_group.background.color = "gray95", 
                    column_labels.background.color = 'gray80', stub.border.width = 0, 
                    data_row.padding.horizontal = 0,
                    data_row.padding = 2)

tab2 = tab2 %>% cols_width(everything()~185)
tab2
## Saving the table

gtsave(tab2, filename = "table1.pdf", path = "./")
gtExtras::gtsave_extra(tab2, filename = "table1.2.png",path = "./")

```

```{mermaid}
%%| fig-width: 6.5
flowchart LR
  A{Number of \nwindows\nin home} --> B{Tuberculosis\ninfection}
  C(Sex) --> B
  A --> D(Household members \nwith tuberculosis)
  D -->B
  H(BCG\nvaccination) --> B
  I(Smoking) --> B
  E(Household density) --> B
  E --> A
  F(Income)--> E
  G(Age) --> F
  C -->F
  G --> B
  F --> A
  
```